# C.elegans Plate Experiment

---

## Exploring the Excel Dataset Structure

**"Review the following Excel file in `./Opdracht 4 - Reproducible Science/4a - CELIQ/CE.LIQ.FLOW.062_Tidydata.xlsx`..."**

The **Parameters** tab is unclear — it's difficult to interpret what each column means. The **Raw Data** column lacks explanation, and the purpose of the coloring (particularly the long pink section) is not described. Other tabs are even more confusing due to minimal labeling and inconsistent formatting.

---

## Importing the Dataset into R

**"Open the file in R using the `{readxl}` package."**

```{r import-data, message=FALSE, warning=FALSE}
library(here)
library(readxl)
library(tidyverse)

# Import data
CElegans_data <- read_excel(
  here::here("Opdracht 4 - Reproducible Science", "4a - CELIQ", "CE.LIQ.FLOW.062_Tidydata.xlsx")
)

# Preview
head(CElegans_data)

```
## Checking and Correcting Data Types

compConcentration was imported as character and was converted to numeric. compName and expType were converted to factors to allow proper grouping and visualization.

```{r inspect data}
# Inspect column types
str(CElegans_data %>% select(RawData, compName, compConcentration))

# Convert data types as needed
CElegans_data$compName <- factor(CElegans_data$compName, levels = unique(CElegans_data$compName))
CElegans_data$expType <- factor(CElegans_data$expType, levels = unique(CElegans_data$expType))
CElegans_data$compConcentration <- parse_number(CElegans_data$compConcentration)

# Confirm conversion
str(CElegans_data %>% select(RawData, compName, compConcentration, expType))
```
## Visualizing Raw Data: Scatterplot
```{r ggplot}
library(ggplot2)
library(tidyverse)


# Plotting the scatterplot with dynamic shape assignment
ggplot(CElegans_data, aes(x = log10(compConcentration), y = RawData)) +
  geom_point(
    aes(color = compName, shape = expType),
    size = 1.5,
    alpha = 0.8,
    position = position_jitter(width = 0.15)
  ) +
  labs(
    title = "C. elegans Response by Compound and Concentration",
    caption = "Raw data colored by compound, shaped by experiment type",
    x = "log10(Compound Concentration) [nM]",
    y = "Raw Data (counts)"
  ) +
  scale_color_brewer(palette = "Dark2") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

## Fixing Axis Label Issues in ggplot
The compConcentration column was originally imported as a character variable, so ggplot2 treated it as a categorical (discrete) variable. This led to disorganized and overlapping axis labels. After converting it to numeric and applying log10(), the axis behaved as continuous, solving the issue.

## Identifying Positive Controls
The positive control for this experiments is controlPositive.

## Identifying Negative Controls
The negative control for this experiment is controlNegative.

## 	Designing a Statistical Analysis Strategy
1) Import and clean the dataset
2) Convert character columns to numeric or factor types
3) Visualize raw data for trends by concentration and compound
4) Normalize RawData against the mean of the controlNegative group
5) Re-plot normalized data to observe relative trends
6) Check normality using Shapiro-Wilk tests
7) Run ANOVA to detect significant differences across groups
8) Apply post hoc tests if ANOVA is significant (e.g., Tukey HSD)
9) Fit dose-response models to estimate IC50 values
10) Compare confidence intervals between IC50 estimates
11) Interpret results for biological and statistical significance

## Normalizing Data Using the Control Group

**"Normalize the data for the controlNegative in such a way that the mean value for controlNegative is exactly equal to 1 and that all other values are expressed as a fraction thereof. Rerun your graphs with the normalized data."**

```{r normalise data}
# Calculate the mean for controlNegative
neg_ctrl <- CElegans_data %>% filter(expType == "controlNegative")
neg_mean <- mean(neg_ctrl$RawData)

# Normalize the RawData
CElegans_data <- CElegans_data %>%
  mutate(Normalized = RawData / neg_mean)

# Preview normalized data
head(CElegans_data %>% select(compName, expType, compConcentration, RawData, Normalized))
```

```{r normalised plot}
ggplot(CElegans_data, aes(x = log10(compConcentration), y = Normalized)) +
  geom_point(
    aes(color = compName, shape = expType),
    size = 1.5,
    alpha = 0.8,
    position = position_jitter(width = 0.15)
  ) +
  labs(
    title = "Normalized C. elegans Response",
    caption = "Data normalized to controlNegative (mean = 1)",
    x = "log10(Compound Concentration) [nM]",
    y = "Normalized Raw Data"
  ) +
  scale_color_brewer(palette = "Dark2") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 	Why Normalize?
This normalization allows for clearer comparison of relative differences between conditions. By setting the control to 1, the effect size of each treatment becomes easier to interpret.
